Confronting Wicked Crypto

### 2021-02-11 articles/tots-ieee/Article.tex
Update TOTS IEEE article, copy in chapter 5

\chapter{EA Debate Arguments}
\label{chap-arguments}

@@

\begin{sidewaysfigure}
  \centering\CaptionFontSize
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.contrib.pdf}
  \caption{Contributing Factors to the EA Debate}
  \label{fig-args-contrib}
\end{sidewaysfigure}

@@

Information security is the remaining central issue in the debate. As described in \mysec{sec-crypto-basics},
\ac{cryptography} plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought
to compromise the cryptographic foundations of encryption; however, this is seen as too risky today.
\mysec{sec-tech-approaches} introduced several alternative technical approaches to \ac{EA}, but experts still emphasize
that current systems cannot securely provide the level of access \acl{LE} asks for \cite{abelson_2015}
\cite{abelson_risks_1997}. Security was an afterthought in early computing and networking designs. The field of
cybersecurity is still equal parts art and science; mandated \ac{EA} would put a heavy burden on a field still finding
its legs.

@@

\begin{figure}[t!]
  \centering\CaptionFontSize
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goingdark.pdf}
  \caption{A ``Going Dark'' Argument Map}
  \label{fig-arg-going-dark}
\end{figure}

@@

\begin{figure}[p!]
  \centering\CaptionFontSize
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goldenage.pdf}
  \caption{A ``Golden Age for Surveillance'' Argument Map}
  \label{fig-arg-golden-age}
\end{figure}

@@

Still, law enforcement's inability to access the tools of \ac{masssurv} cannot be taken for granted. It will likely
change---they already have access to mobile devices. A more conservative criticism of the ``Golden Age'' argument rests
in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement while
thwarting \ac{masssurv} subvert most of the ``Golden Age'' argument. \ac{EA} proposals with these aims will be explored
in \mysec{sec-ea-types}.

@@

\begin{figure}[t!]
  \centering\CaptionFontSize
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.fallacies.pdf}
  \caption{Fallacious Arguments of the EA Debate}
  \label{fig-arg-fallacies}
\end{figure}

@@

The last four maps framed the debate, presented leading arguments, and identified fallacious arguments. The next map
explores potential solutions. \myfig{fig-arg-measures} divides potential solutions into the categories of current
capabilities, legal measures, and \ac{EA}. Current capabilities can be implemented today, and include maintenance of the
status quo and increasing investment in current programs. Legal measures require a change in the text or application of
the law, but do not require fundamentally new technical capabilities. These include compelling passwords, which requires
legal clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
non-\ac{EA} approaches were first introduced in \mysec{sec-alt-approaches}.) In this section, \ac{EA} is compared to its
alternatives. \ac{EA} variations are compared in the next map.

@@

\begin{sidewaysfigure}
  \centering\CaptionFontSize
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.measures.pdf}
  \caption{EA and its Alternatives}
  \label{fig-arg-measures}
\end{sidewaysfigure}

@@

\begin{sidewaysfigure}[p!]
  \centering\CaptionFontSize
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.classes.pdf}
  \caption{Classes of EA}
  \label{fig-arg-classes}
\end{sidewaysfigure}

@@

Device encryption, particularly in mobile phones, may actually pose the smallest impediment to investigation of any
encryption technology. This fact undercuts the argument for device key escrow. The ``Mass Extraction'' report clearly
describes the current state of affairs: device manufacturers are losing the exploit arms race, private companies are
commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory strategy is
leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs reinforces the
lessons of \mychap{chap-policy}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has
led to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This
is because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
accessing devices again, this issue will immediately resurface. This is likely why law enforcement continues to push for
\ac{EA} despite present circumstances. Additionally, a formalized process may reduce government abuse, which threatens
privacy and civil liberties more systematically than criminal abuse.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% SETUP %
\documentclass[conference]{IEEEtran}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

%%%%%%%%%%% Non-standard packages
\usepackage[hyphens]{url}
\usepackage[acronym,toc,nonumberlist]{glossaries}
\usepackage[shortcuts=ac]{glossaries-extra}
\usepackage{rotating}
\usepackage{csquotes}

%%%%%%%%%%% Commands
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\tt#1{\mbox{\texttt{#1}}}

\newcommand*{\thead}[1]{\multicolumn{1}{c}{\bfseries #1}}

%% Mine
\def\tt#1{\mbox{\texttt{#1}}}
\def\bb#1{\mbox{\textbf{#1}}}
\def\ii#1{\mbox{\textit{#1}}}
\newcommand{\myfig}[1]{Figure~\ref{#1}}
\newcommand{\mysec}[1]{Section~\ref{#1}}

%%%%%%%%%%% Inputs
\input{../../Thesis-Glossary}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TITLE %

\title{Confronting Wicked Crypto}

\author{\IEEEauthorblockN{Kevin Kredit}
\IEEEauthorblockA{\textit{GVSU ACS Department} \\
Grand Rapids, Michigan \\
Email: k.kredit.us@ieee.org}
\and
% FIXME: Prof. Kalafut, update as appropriate
% Is there a way to indicate that you are faculty and I am not? Label yourself with PhD, perhaps?
\IEEEauthorblockN{Andrew Kalafut, Ph.D.}
\IEEEauthorblockA{\textit{GVSU ACS Department} \\
Grand Rapids, Michigan \\
Email: kalafuta@gvsu.edu}
}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% ABSTRACT %

\begin{abstract}

% NOTE: I think we should switch terminology to "Lawful access" for the paper (as opposed to EA)

% NOTE: The first paragraph of the abstract was copied. The second I tweaked.

Public debate has resumed on the topic of lawful access (LA) to encrypted data, which refers to alternative means of
decryption intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at
the end of the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
technology is a wicked problem that has and will continue to evade a permanent solution. As policymakers consider next
steps, it is necessary that the technical community remain engaged. Although any LA framework would increase risk, the
magnitude of that increase varies greatly with the quality of the technical and regulatory approach. Furthermore, if one
considers hard-line legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed
LA may reduce risk overall.

The root of the conflict lies in cryptography's dual role as an enabler of unprecedented privacy and a
cornerstone of security. The emergence of strong encryption incited the first crypto war, and its proliferation is
causing the second. In order to make progress in the debate, this paper frames encryption and LA as a wicked problem and
analyze the debate using argument maps.

\end{abstract}

\begin{IEEEkeywords}
encryption, exceptional access, policy, wicked problems
\end{IEEEkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% PAPER %

\section{Introduction}
\label{sec-intro}

% TODO: add actual intro, not just chapter intro. Take some from paper intro, but also briefly discuss wicked problems

@@

\begin{sidewaysfigure*}
  \centering
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.contrib.pdf}
  \caption{Contributing Factors to the EA Debate}
  \label{fig-args-contrib}
\end{sidewaysfigure*}

@@

Information security is the remaining central issue in the debate. As described in \mysec{sec-intro}, \ac{cryptography}
plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought to compromise the
cryptographic foundations of encryption; however, this is seen as too risky today. \mysec{sec-intro} introduced several
alternative technical approaches to \ac{EA}, but experts still emphasize that current systems cannot securely provide
the level of access \acl{LE} asks for \cite{abelson_2015} \cite{abelson_risks_1997}. Security was an afterthought in
early computing and networking designs. The field of cybersecurity is still equal parts art and science; mandated
\ac{EA} would put a heavy burden on a field still finding its legs.

% TODO: fix \mysec{sec-intro} reference in above paragraph
% TODO: fix \mysec{sec-intro} reference in above paragraph (there are two instances)

@@

\begin{figure*}[t!]
  \centering
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goingdark.pdf}
  \caption{A ``Going Dark'' Argument Map}
  \label{fig-arg-going-dark}
\end{figure*}

@@

\begin{figure*}[p!]
  \centering
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.goldenage.pdf}
  \caption{A ``Golden Age for Surveillance'' Argument Map}
  \label{fig-arg-golden-age}
\end{figure*}

@@

Still, law enforcement's inability to access the tools of \ac{masssurv} cannot be taken for granted. It will likely
change---they already have access to mobile devices. A more conservative criticism of the ``Golden Age'' argument rests
in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement while
thwarting \ac{masssurv} subvert most of the ``Golden Age'' argument. \ac{EA} proposals with these aims will be explored
in \mysec{sec-intro}.

% TODO: fix \mysec{sec-intro} reference in above paragraph

@@

\begin{figure*}[t!]
  \centering
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.fallacies.pdf}
  \caption{Fallacious Arguments of the EA Debate}
  \label{fig-arg-fallacies}
\end{figure*}

@@

The last four maps framed the debate, presented leading arguments, and identified fallacious arguments. The next map
explores potential solutions. \myfig{fig-arg-measures} divides potential solutions into the categories of current
capabilities, legal measures, and \ac{EA}. Current capabilities can be implemented today, and include maintenance of the
status quo and increasing investment in current programs. Legal measures require a change in the text or application of
the law, but do not require fundamentally new technical capabilities. These include compelling passwords, which requires
legal clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
non-\ac{EA} approaches were first introduced in \mysec{sec-intro}.) In this section, \ac{EA} is compared to its
alternatives. \ac{EA} variations are compared in the next map.

% TODO: fix \mysec{sec-intro} reference in above paragraph

@@

\begin{sidewaysfigure*}
  \centering
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.measures.pdf}
  \caption{EA and its Alternatives}
  \label{fig-arg-measures}
\end{sidewaysfigure*}

@@

\begin{sidewaysfigure*}[p!]
  \centering
  \includegraphics[width=\linewidth]{arguments/build/comprehensive.classes.pdf}
  \caption{Classes of EA}
  \label{fig-arg-classes}
\end{sidewaysfigure*}

@@

Device encryption, particularly in mobile phones, may actually pose the smallest impediment to investigation of any
encryption technology. This fact undercuts the argument for device key escrow. The ``Mass Extraction'' report clearly
describes the current state of affairs: device manufacturers are losing the exploit arms race, private companies are
commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory strategy is
leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs reinforces the
lessons of \mysec{sec-intro}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has led
to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This is
because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
accessing devices again, this issue will immediately resurface. This is likely why law enforcement continues to push for
\ac{EA} despite present circumstances. Additionally, a formalized process may reduce government abuse, which threatens
privacy and civil liberties more systematically than criminal abuse.

% TODO: fix \mysec{sec-intro} reference in above paragraph

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BIBLIOGRAPHY %

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,Article} % IEEEfull or IEEEabrv

\end{document}
